GPU Acceleration of Fixed Complexity Sphere Decoder for Large MIMO Uplink Systems

### 2014-11-05 parallel_acceleration_of_FCSD_based_on_cuda/draft.tex
solve the crush problem and revise the report

\paragraph{}General purpose Graphic processing Units Computing (GPGPU) has become a common trend in industry and research in areas that need massive computations such as computer vision\cite{fung2008using}, computer games\cite{blewitt2013applicability}, signal processing\cite{van2011accelerating}\cite{6671435} and finance\cite{grauer2013accelerating}. Compared to CPU, the GPU can naturally cope with computational intensive tasks, since most of the hardware resource are allocated for data processing and parallel computation, rather than caching and flow control. In recent years GPU computing has provided significant advantages in data processing speeds over CPU computations\cite{nvidia2008programming}. Computed unified Device Architecture (CUDA) is a programming model for GPGPU that was proposed by Nvidia and is becoming the dominant framework for GPGPU programming. It is based on popular programming languages(C, C++, Java, Python, Fortran, Direct Compute), such that the application developers in different areas do not need to learn a new programming language or get to know the hardware details. CUDA employs the Single Instruction Multiple Threads (SIMT) programming model, enables a GPU to implement parallel data processing, and has become a parallel computing platform for wide community of different applications.

@@

\paragraph{}In Figure \ref{figure2} we use the GPU-GeForce GTX 760 as an example to show the microarchitecture of  GPU. The GPU works as a coprocessor system, consisting of Stream Multiprocessors (SM). The GeForce GTX 760 has 6 SM that perform the multiply-add arithmetic operation. Furthermore, each SM has the special functional units(SFU) that can perform more complex arithmetics such as trigonometric functions and reciprocal square root. The shared memory, also called on-chip memory located on SM and has low latency and limited capacity. Global memory, which is off-chip memory of the GPU has long latency and large capacity. Another two types of the memory, texture memory and constant memory, are located off-chip, but cached. Therefore the loading speed of the data located in these two type of memory space is much faster than that from global memory. Constant memory is read only and with latency approximately as small as the registers.  The variables invoked in $\mathit{kernel}$ functions can be stored in either of the these four types of memories. Registers are allocated dynamically to threads.

@@

\paragraph{}As mention in section \ref{programming model}, our implementation is based on heterogeneous programming principle, the flow control and logical operations are reside at the host side, the compute intensive works is implemented at device side. Although the FCSD preprocessing has a serial nature and can not be vectorized, the computational consumptive operations can be performed by the CUDA Basic Linear Algebra Subprograms(cuBLAS) and corresponding GPU program implementations, as shown in table \ref{table1}.
\begin{table}[htb]
\centering
\begin{tabular}{|c|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
\hline
Operations & complex matrix-matrix multiplication & complex matrix-vector multiplication & matrix inverse & cholesky factorization \\
\hline
GPU subroutine & $\mathit{cublasCgemm}$  & $\mathit{cublasCgemv}$ & $\mathit{GJE}$ & $\mathit{chol}$\\
\hline
Formula   &  $\mathbf{H}^{H}\mathbf{H}$ & $(\mathbf{H}^{H}\mathbf{H})^{-1}\mathbf{H}^{H}y$ & $\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1} $ & $\mathbf{H}^{H}\mathbf{H}=\mathbf{R}^{H}\mathbf{R}$ 
\\
\hline
\end{tabular}
\caption{Parallel Computation Operation of FCSD Preprocessing}
\label{table1}
\end{table}

@@

\paragraph{}GPU is computation intensive and only has small cache for special memory space (texture, constant). Thus the memory reading$\&$writing (R/W) strategy of GPU is the crutial factor that influences the application performance. Different memory types has different valid scope, local memory which is invoked in kernel function is valid only in one thread, share memory is valid in one block, constant, texture and global memory are valid both at host side and device side. On the other hand different from global memory (2047 MBytes in GeForce GTX 760), the constant and share memory are all capacity limited (only $65536$ bytes totally and $49152$ bytes per block in GeForce GTX 760). Thus the occupancy of memory is one determinant factor needed to be considered. 

@@

 \paragraph{}In table \ref{amount of R/W operation} we present the amount of R/W operations and the storage that needed for different data variables in FCSD path searching. The amount of the R/W operation is calculated based on the (\ref{path searching}) and (\ref{Eu metric}), for complex values ($\mathbf{R}$ $\mathbf{\hat{s}}$, $\mathit{s_{pm}}$, $\mathit{s_{kernel}}$ ) are stored using CUDA build in variable data type $\mathit{cuComplex}$ which counts 8 bytes. Floating point data set $E_{u}$ is stored using variable type $\mathit{float}$, which counts for 4 bytes, $\mathit{s_{index}}$ is the integer list which is stored using variable type $\mathit{int}$, which counts 1 byte.\\

@@

 \renewcommand{\arraystretch}{1.3}

@@

\begin{tabular}{|p{3cm}|p{3cm}|l|p{3cm}|l|l|l|}
\hline
data & $\mathbf{R}$ & $\mathbf{\hat{s}}$ & $\mathit{s_{pm}}$ & $\mathit{s_{index}}$ & $\mathit{E_{u}}$ & $\mathit{s_{kernel}}$ \\
\hline
R/W operations/thread & $(\frac{3}{2}(N_{t})^{2}-\frac{N_{t}}{2}+\rho-\rho^{2})$ & $(\frac{2(N_{t})^{2}+2N_{t}-\rho^{2}-\rho}{2})$ & $N_{t}$ & $\rho$ &$N_{t}$ & $N_{t}$ \\
\hline
Storage/bytes  & $4(N_{t}+1)N_{t}$ & $8N_{t}$ & $8M^{\rho}N_{t}$ & $N_{t}M^{\rho}$ & $4N_{t}M^{\rho}$ & $8N_{t}$\\
\hline

\end{tabular} 
 \label{amount of R/W operation} 
 \caption{Amount of R/W Operations and Storage Requirement of Different Data Set}
\end{table} 
\subsubsection{Memory Accesss Pattern}
  \paragraph{}When considering large MIMO array, the large amount of R/W operations determine that the memory bandwidth is the major bottleneck of the performance,  although the theoretical bandwidth of off-chip memory is extremely high (in GeForce GTX 760 the bandwidth of off-chip memory is around $192 GB/s$), however the effect memory bandwidth can only be achieved by proper R/W strategy.\\
  \paragraph{}Global memory is implemented by dynamic random access memory (DRAM) whose R/W is extremely slow process, modern DRAM R/W operations are performed by access a pitch of consecutive memory location, as mentioned in section (\ref{programming model}), the warp is the basic schedule unit of device, with one instruction recall one warp of threads that execute at the same time, the most favorable global memory access pattern is to keep one warp of threads that access consecutive memory as much as possible. For the GeForce GTX 760, the DRAM will access a consecutive memory address that has 128 bytes length, In Figure \ref{coalesce global memory} we take the global access pattern of the $\mathit{s_{pm}}$ as an example to illustrate how we coalesce the the global memory.
  \begin{figure}[tb]
\centering
\includegraphics[scale=0.63]{coalescing_global_memory.eps}
\caption{Global Memory Access Pattern of $\mathit{s_{pm}}$}
\label{coalesce global memory}
\end{figure}
We store $\mathit{s_{pm}}$ in one pitch of linear global memory, with each row store one $N_{t}=16$ complex symbol vector solution candidate of $\mathbf{\hat{s}}^{F}$ in (\ref{formula 5}), totally there are $M^{\rho}=4096$ rows, this matrix is stored in column major. As mentioned before one path searching is performed by one thread, thus the address two consecutive symbol operated in one thread is $4096$, however for consecutive threads the address they reach is continues as shown in Figure \ref{coalesce global memory}, each 32 threads in one warp can make R/W operation to a pitch of continues memory, which is $32\times 8=256$ bytes, counts for two times of DRAM access.

 \paragraph{}Another available resource for memory bandwidth increase is to make use of share memory, for all the transient variables that used for accumulator in (\ref{path searching}) and (\ref{Eu metric}), because of the small storage consumption and large amount of R/W operations.   
\subsubsection{Data Transfer Minimization}   
\paragraph{}The data transfer between the host and the device side is much more slower than the inner device data transferring, therefore, in our application we minimize the data transferring times, we make data transfer only once for propagation channel matrix $\mathbf{H}$, received symbol vector $\mathbf{y}$ in (\ref{formula 1}) as well as the sub symbol vector index list $\mathit{s_{index}}$ at FE stage, we integrate the post processing (\ref{Eu metric}) into path searching kernel function to avoid the transferring of $E_{u}$ in (\ref{Eu metric}).  
\subsection{Going to Larger MIMO System}
  \paragraph{}Consider the uplink $32\times 32$ MIMO system with $16$QAM modulation scheme, the path number that need to be searched is $M^{\rho}=16^{5}=1048576$. The maximum number of threads that can be execute in $\mathit{kernel}$ function of the GeForce GTX 760 is 12288, therefore we need to consider execute the $\mathit{kernel}$ function in loop, in order to avoid device synchronization latency, we synchronize and return the control to host side in order to reduce the synchronization expense. Under this condition we use 12 linear blocks and each block has 1024 threads, we execute the $\mathit{kernel}$ function for 86 times and get the final solution,  when it comes to the large MIMO systems and large signal constellations such as $36\times 36$ $16$QAM, $144\times 144$ $4$QAM, $20\times 20$ $64$QAM, the global memory that need to used to store $s_{pm}$ in section (\ref{data preparation}) will out of bound (2 GB), multiple kernels with each one deal with a small subset of symbol vector candidates and kernel concurrency\cite{nvidia2008programming} can be applied to solve this problem.
\section{Simulation Results}\label{simulation}
\subsection{Environment}
\subsubsection{Device}
\emph{Graphic Processing Units}: GeForce GTX 760, GPU clock rate: 1.08 GHz , memory clock rate: 3 GHz , number of SM: 6, maximum number of threads per stream multiprocessor: 2048, total amount of global memory: 2 GB, total amount of share memory per block: 49 KB, total amount of constant memory: 66 KB. \\
\emph{Central Processing Units-Modigliani}: Intel Core i5-4th generation, 4 cores 3.20 GHz CPU clock rate, 8 GB RAM\\
\emph{Central Processing Units-Monet} : Intel Core i7-3rd generation, 12 cores 3.20 GHz CPU clock rate, 32 GB RAM
\subsubsection{Software}
CUDA Driver Version / Runtime Version      6.5 / 6.0\\
Nsight Eclipse Version 6.5
\subsection{Performance and Evaluation}
\begin{table}[htb]
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multirow{2}{*}{ Array size} & \multicolumn{3}{|c|}{Time/s} & \multicolumn{2}{|c|}{Speedup}\\
\cline{2-6}
&GeForce GTX 760 & Modigliani & Monet &  GTX 760/Modigliani  &  GTX 760/Monet \\
\hline
$8\times 8$& 1.88& 0.10&0.11 & 0.05& 0.06\\
\hline
$16\times 16$&6.12 & 0.92&0.98& 0.15& 0.16\\
\hline
$32\times 32$&38.32 & 31.27& 34.60& 0.82& 0.90\\
\hline
$48\times 48$&113.80& 262.00& 285.14& 2.30& 2.51\\
\hline
$64\times 64$& 537.41&1753.30&1940.10&3.26& 3.61 \\
\hline
$72\times 72 $&2294.32&8727.69 &9641.80 &3.80 &4.20 \\
\hline
\end{tabular}
\caption{speedup performance of different MIMO systems using 4 QAM}
\label{speedup1}
\end{table}

@@

\paragraph{} We made simulation to different sizes MIMO systems and signal constellation and comparison is made between the Geforce GTX 760 and two kinds of CPU, the speedup performance is presented in table \ref{speedup1}, \ref{speedup2} and \ref{speedup3}. All the operation times we test is based on $1000$ channel realizations, the transmitted data is generated randomly and modulated by $M$ quadrature amplitude modulation (QAM) scheme.
\paragraph{} In table \ref{speedup1}, for small MIMO systems $8\times 8$ and $16\times 16$, the speed performance of GPU implementation is even worse than that of CPU. The number of path of these antenna configurations that need to be searched is relatively small ($16$ for $8\times 8$ system and $64$ for $16\times 16$ system), the CPU implementation has already had a good performance, there is a short time for GPU to "warm up", for host-device data transmission, launch kernel function and host device synchronization. Furthermore, for small number of threads, the memory R/W latency is not well hided. For the $32\times 32$ systems, the speed performance of GPU and CPU implementation is close, when comes to larger MIMO systems $48\times 48$, $64\times 64$, $72\times 72$, a large parallelism is achieved so that the GPU begin to display its data processing power, the speedup can be from $2.50$ to $4.20$. 
\paragraph{}In table \ref{speedup2} and \ref{speedup3}, $16$QAM and $64$QAM are considered, the speedup of GPU implementation is reached from a small MIMO size, For $16\times 16$ system using $16$QAM $1.18$ speedup over Modigliani and $1.57$ speedup over Monet. For $8\times 8$, $64$QAM, $0.87$ speedup over Modigliani and $1.09$ speedup over Monet.  For lager signal constellation, the massive parallelism can be reached for relatively small MIMO systems.
\paragraph{}It can be observed from table \ref{speedup1}, \ref{speedup2} and \ref{speedup3}, GPU implementation shows its speed advantage over CPU implementation for increasing size of MIMO systems and signal constellations, indicating this GPU implementation's suitability for large MIMO system and large signal constellation.
\begin{figure}[htb]
\centering
\includegraphics[scale=0.65]{BER_curves.eps}
\caption{BER performance of different sizes of MIMO systems and modulation scheme by GPU}
\label{BER curve}
\end{figure}
\paragraph{}Figure \ref{BER curve} shows the Monte Carlo simulation results of bit error rate (BER) performance for different MIMO systems and constellation sizes, we consider uncoded spatial multiplexing MIMO systems, the transmitted binary sequence is generated randomly and  mutually independent, modulated by $M$-QAM modulation scheme, the results have been obtained using $100000$ channel realizations and $500$ symbol errors accumulated. 
\section{Conclusion}\label{conclusion}
\paragraph{}This paper presents the GPU implementation of fixed complexity sphere decoder for a large MIMO system, which is a time consuming task for traditional computation platform. In order to exploit the computational capability of GPU in the simulation of large MIMO systems, the utilization of heterogeneous programming model, thread resource limitation and memory configuration are considered in a comprehensive way. The simulation shows considerable acceleration of GPU implementation of FCSD can be reached for large MIMO systems and large signal constellation sizes while the same BER performance can be obtained. In addition, the GPU and CPU can work independently, the data processing power of GPU computing is an extra bonus of the traditional simulation platform. This shows the potential of GPU computing for computational consumptive simulation in large MIMO area. The future work contains taking into consider the kernel execution and data transferring concurrence in order to optimize memory allocation strategy for large MIMO systems.       
\newpage
% conference papers do not normally have an appendix

>>>>

\paragraph{}General purpose Graphic processing Units Computing (GPGPU) has become a common trend in industry and research in areas that need massive computations, such as computer vision\cite{fung2008using}, computer games\cite{blewitt2013applicability}, signal processing\cite{van2011accelerating}\cite{6671435} and finance\cite{grauer2013accelerating}. Compared to CPU, the GPU can naturally cope with computational intensive tasks, since most of the hardware resource are allocated for data processing and parallel computation, rather than caching and flow control. In recent years GPU computing has provided significant advantages in data processing speeds over CPU computations\cite{nvidia2008programming}. In 2011, the NVIDIA GPUs have become the fastest supercomputer in the world\cite{cook2013cuda},  Computed unified Device Architecture (CUDA) is a programming model for GPGPU that was proposed by NIVIDIA and has become the dominant framework for GPGPU programming. It is based on popular programming languages(C, C++, Java, Python, Fortran, Direct Compute), such that the application developers in different areas do not need to learn a new programming language or get to know the hardware details. CUDA employs the Single Instruction Multiple Threads (SIMT) programming model, enables a GPU to implement parallel data processing, and has become a parallel computing platform for wide community of different applications.

@@

\paragraph{}In Figure \ref{figure2} we use the GPU-GeForce GTX 760 as an example to show the microarchitecture of  GPU. The GPU works as a coprocessor system, consisting of Stream Multiprocessors (SM). The GeForce GTX 760 has 6 SM that perform the multiply-add arithmetic operation. Furthermore, each SM has the special functional units(SFU) that can perform more complex arithmetics such as trigonometric functions and reciprocal square root. The shared memory, also called on-chip memory located on SM and has low latency and limited capacity. Global memory, which is off-chip memory of the GPU has long latency and large capacity. Another two types of the memory, texture memory and constant memory, are located off-chip, but cached. Therefore the loading speed of the data located in these two type of memory space is much faster than that from global memory. Constant memory is read only and with latency approximately as small as the registers.  The variables invoked in $\mathit{kernel}$ functions can be stored in either of the these four types of memories. Registers are allocated dynamically and privately to threads, the access speed of registers is the faster than all the other four types of memories.

@@

\paragraph{}As mention in section \ref{programming model}, our implementation is based on heterogeneous programming principle-achieving performance with a mix of CPU and GPU technology, the flow control and logical operations are reside at the host side, the compute intensive works is implemented at device side. Although the FCSD preprocessing has a serial nature and can not be vectorized, the computational consumptive operations can be performed by the CUDA Basic Linear Algebra Subprograms(cuBLAS) and corresponding GPU program implementations, as shown in table \ref{table1}.
\begin{table}[htb]
\centering
\begin{tabular}{|c|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
\hline
Operations & complex matrix-matrix multiplication & complex matrix-vector multiplication & matrix inverse & cholesky factorization \\
\hline
GPU subroutine & $\mathit{cublasCgemm}$  & $\mathit{cublasCgemv}$ & $\mathit{GJE}$ & $\mathit{chol}$\\
\hline
Formula   &  $\mathbf{H}^{H}\mathbf{H}$ & $(\mathbf{H}^{H}\mathbf{H})^{-1}\mathbf{H}^{H}y$ & $\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1} $ & $\mathbf{H}^{H}\mathbf{H}=\mathbf{R}^{H}\mathbf{R}$ 
\\
\hline
\end{tabular}
\caption{Parallel Computation Operation of FCSD Preprocessing}
\label{table1}
\end{table}

@@

\paragraph{}Unlike CPUs that have a flat memory model meaning the CPU core can access any memory location without restriction, GPUs are computation intensive and only have small cache for special memory space (texture, constant). Thus the memory reading$\&$writing (R/W) strategy of GPU is the crutial factor that influences the application performance. Different memory types has different valid scope, local memory which is invoked in kernel function is valid only in one thread, share memory is valid in one block, constant, texture and global memory are valid both at host side and device side. On the other hand different from global memory (2047 MBytes in GeForce GTX 760), the constant and share memory are all capacity limited (only $65536$ bytes totally and $49152$ bytes per block in GeForce GTX 760). Thus the occupancy of memory is one determinant factor needed to be considered. 

@@

 \paragraph{}In table \ref{amountofR/Woperations}  we present the amount of R/W operations and the storage that needed for different data variables in FCSD path searching. The amount of the R/W operation is calculated based on the (\ref{path searching}) and (\ref{Eu metric}), for complex values ($\mathbf{R}$ $\mathbf{\hat{s}}$, $\mathit{s_{pm}}$, $\mathit{s_{kernel}}$ ) are stored using CUDA build in variable data type $\mathit{cuComplex}$ which counts 8 bytes. Floating point data set $E_{u}$ is stored using variable type $\mathit{float}$, which counts for 4 bytes, $\mathit{s_{index}}$ is the integer list which is stored using variable type $\mathit{int}$, which counts 1 byte.\\

@@

@@

\begin{tabular}{|p{3cm}|p{3cm}|l|p{3cm}|l|l|l|}
data & $\mathbf{R}$ & $\mathbf{\hat{s}}$ & $\mathit{s_{pm}}$ & $\mathit{s_{index}}$ & $\mathit{E_{u}}$ & $\mathit{s_{kernel}}$ \\
R/W operations/thread & $(\frac{3}{2}(N_{t})^{2}-\frac{N_{t}}{2}+\rho-\rho^{2})$ & $(\frac{2(N_{t})^{2}+2N_{t}-\rho^{2}-\rho}{2})$ & $N_{t}$ & $\rho$ &$N_{t}$ & $N_{t}$ \\
Storage/bytes  & $4(N_{t}+1)N_{t}$ & $8N_{t}$ & $8M^{\rho}N_{t}$ & $N_{t}M^{\rho}$ & $4N_{t}M^{\rho}$ & $8N_{t}$\\
\end{tabular} 
 \caption{Amount of R/W Operations and Storage Requirement of Different Data Set}
  \label{amountofR/Woperations} 
\end{table}  
\subsubsection{Memory Accesss Pattern}
  \paragraph{}When considering large MIMO array, the large amount of R/W operations determine that the memory bandwidth is the major bottleneck of the performance,  although the theoretical bandwidth of off-chip memory is extremely high (in GeForce GTX 760 the bandwidth of off-chip memory is around $192 GB/s$), however the effect memory bandwidth can only be achieved by proper R/W strategy.\\
  \paragraph{}Global memory is implemented by dynamic random access memory (DRAM) whose R/W is extremely slow process, modern DRAM R/W operations are performed by access a pitch of consecutive memory location, as mentioned in section (\ref{programming model}), the warp is the basic schedule unit of device, with one instruction recall one warp of threads that execute at the same time, the most favorable global memory access pattern is to keep one warp of threads that access consecutive memory as much as possible. For the GeForce GTX 760, the DRAM will access a consecutive memory address that has 128 bytes length, In Figure \ref{coalesce global memory} we take the global access pattern of the $\mathit{s_{pm}}$ as an example to illustrate how we coalesce the the global memory.
  \begin{figure}[tb]
\includegraphics[scale=0.63]{coalescing_global_memory.eps}
\caption{Global Memory Access Pattern of $\mathit{s_{pm}}$}
\label{coalesce global memory}
We store $\mathit{s_{pm}}$ in one pitch of linear global memory, with each row store one $N_{t}=16$ complex symbol vector solution candidate of $\mathbf{\hat{s}}^{F}$ in (\ref{formula 5}), totally there are $M^{\rho}=4096$ rows, this matrix is stored in column major. As mentioned before one path searching is performed by one thread, thus the address two consecutive symbol operated in one thread is $4096$, however for consecutive threads the address they reach is continues as shown in Figure \ref{coalesce global memory}, each 32 threads in one warp can make R/W operation to a pitch of continues memory, which is $32\times 8=256$ bytes, counts for two times of DRAM access.

 \paragraph{}As we mentioned in section (\ref{programming model}), the registers are allocated to the threads dynamically and have much faster access speed than that of global memory. Therefore, for all the transient data that used for accumulator in (\ref{path searching}), (\ref{Eu metric}), and the symbol vector candidate during the path searching process of each thread, are stored temporarily in registers to hide the latency of R/W of global memory.
\subsubsection{Data Transfer Minimization}   
\paragraph{}The data transfer between the host and the device side is much more slower than the inner device data transferring, therefore, in our application we minimize the data transferring times, we make data transfer only once for propagation channel matrix $\mathbf{H}$, received symbol vector $\mathbf{y}$ in (\ref{formula 1}) as well as the sub symbol vector index list $\mathit{s_{index}}$ at FE stage. Loop fusion is applied to post processing stage, we integrate the post processing (\ref{Eu metric}) into path searching kernel function to reduce the iteration times per thread and avoid the transferring of $E_{u}$ in (\ref{Eu metric}).  
\subsection{Going to Larger MIMO System}
  \paragraph{}Consider the uplink $32\times 32$ MIMO system with $16$QAM modulation scheme, the path number that need to be searched is $M^{\rho}=16^{5}=1048576$. The maximum number of threads that can be execute in $\mathit{kernel}$ function of the GeForce GTX 760 is 12288, therefore we need to consider execute the $\mathit{kernel}$ function in loop, in order to avoid device synchronization latency, we synchronize and return the control to host side in order to reduce the synchronization expense. Under this condition we use 12 linear blocks and each block has 1024 threads, we execute the $\mathit{kernel}$ function for 86 times and get the final solution,  when it comes to the large MIMO systems and large signal constellations such as $36\times 36$ $16$QAM, $144\times 144$ $4$QAM, $20\times 20$ $64$QAM, the global memory that need to used to store $s_{pm}$ in section (\ref{data preparation}) will out of bound (2 GB), multiple kernels with each one deal with a small subset of symbol vector candidates and kernel concurrency\cite{nvidia2008programming} can be applied to solve this problem.
\section{Simulation Results}\label{simulation}
\subsection{Environment}
\subsubsection{Device}
\emph{Graphic Processing Units}: GeForce GTX 760, GPU clock rate: 1.08 GHz , memory clock rate: 3 GHz , number of SM: 6, maximum number of threads per stream multiprocessor: 2048, total amount of global memory: 2 GB, total amount of share memory per block: 49 KB, total amount of constant memory: 66 KB. \\
\emph{Central Processing Units-Modigliani}: Intel Core i5-4th generation, 4 cores 3.20 GHz CPU clock rate, 8 GB RAM\\
\emph{Central Processing Units-Monet} : Intel Core i7-3rd generation, 12 cores 3.20 GHz CPU clock rate, 32 GB RAM
\subsubsection{Software}
CUDA Driver Version / Runtime Version      6.5 / 6.0\\
Nsight Eclipse Version 6.5
\subsection{Performance and Evaluation}
\begin{table}[htb]
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multirow{2}{*}{ Array size} & \multicolumn{3}{|c|}{Time/s} & \multicolumn{2}{|c|}{Speedup}\\
\cline{2-6}
&GeForce GTX 760 & Modigliani & Monet &  GTX 760/Modigliani  &  GTX 760/Monet \\
\hline
$8\times 8$& 1.88& 0.10&0.11 & 0.05& 0.06\\
\hline
$16\times 16$&6.12 & 0.92&0.98& 0.15& 0.16\\
\hline
$32\times 32$&38.32 & 31.27& 34.60& 0.82& 0.90\\
\hline
$48\times 48$&113.80& 262.00& 285.14& 2.30& 2.51\\
\hline
$64\times 64$& 537.41&1753.30&1940.10&3.26& 3.61 \\
\hline
$72\times 72 $&2294.32&8727.69 &9641.80 &3.80 &4.20 \\
\hline
\end{tabular}
\caption{speedup performance of different MIMO systems using 4 QAM}
\label{speedup1}
\end{table}

@@

\paragraph{} We made simulation to different sizes MIMO systems and signal constellation and comparison is made between the Geforce GTX 760 and two kinds of CPU, the speedup performance is presented in table \ref{speedup1}, \ref{speedup2} and \ref{speedup3}. All the operation times we test is based on $1000$ channel realizations, the transmitted data is generated randomly and modulated by $M$ quadrature amplitude modulation (QAM) scheme.
\paragraph{} In table \ref{speedup1}, for small MIMO systems $8\times 8$ and $16\times 16$, the speed performance of GPU implementation is even worse than that of CPU. The number of path of these antenna configurations that need to be searched is relatively small ($16$ for $8\times 8$ system and $64$ for $16\times 16$ system), the CPU implementation has already had a good performance, the CPUs are designed for serial code execution with the special hardwares (branch prediction units, multiple caches, etc) and have extremely good performance at it, however the GPUs can achieve a good performance only when they are utilized in parallel manner, there is a short time for GPU to "warm up", for host-device data transmission, launch kernel function and host device synchronization. Furthermore, for small number of threads, the memory R/W latency is not well hided. For the $32\times 32$ systems, the speed performance of GPU and CPU implementation is close, when comes to larger MIMO systems $48\times 48$, $64\times 64$, $72\times 72$, a large parallelism is achieved so that the GPU begin to display its data processing power, the speedup can be from $2.50$ to $4.20$. 
\paragraph{}In table \ref{speedup2} and \ref{speedup3}, $16$QAM and $64$QAM are considered, the speedup of GPU implementation is reached from a small MIMO size, For $16\times 16$ system using $16$QAM $1.18$ speedup over Modigliani and $1.57$ speedup over Monet. For $8\times 8$, $64$QAM, $0.87$ speedup over Modigliani and $1.09$ speedup over Monet.  For lager signal constellation, the massive parallelism can be reached for relatively small MIMO systems.
\paragraph{}It can be observed from table \ref{speedup1}, \ref{speedup2} and \ref{speedup3}, GPU implementation shows its speed advantage over CPU implementation for increasing size of MIMO systems and signal constellations, indicating this GPU implementation's suitability for large MIMO system and large signal constellation.
\includegraphics[scale=0.65]{BER_curves.eps}
\caption{BER performance of different sizes of MIMO systems and modulation scheme by GPU}
\label{BER curve}
\paragraph{}Figure \ref{BER curve} shows the Monte Carlo simulation results of bit error rate (BER) performance for different MIMO systems and constellation sizes, we consider uncoded spatial multiplexing MIMO systems, the transmitted binary sequence is generated randomly and  mutually independent, modulated by $M$-QAM modulation scheme, the results have been obtained using $100000$ channel realizations and $500$ symbol errors accumulated. 
\section{Conclusion}\label{conclusion}
\paragraph{}This paper presents the GPU implementation of fixed complexity sphere decoder for a large MIMO system, which is a time consuming task for traditional computation platform. In order to exploit the computational capability of GPU in the simulation of large MIMO systems, the utilization of heterogeneous programming model, thread resource limitation and memory configuration are considered in a comprehensive way. The simulation shows considerable acceleration of GPU implementation of FCSD can be reached for large MIMO systems and large signal constellation sizes while the same BER performance can be obtained. In addition, the GPU and CPU can work independently, the data processing power of GPU computing is an extra bonus of the traditional simulation platform. This shows the potential of GPU computing for computational consumptive simulation in large MIMO area. The future work contains taking into consider the kernel execution and data transferring concurrence in order to optimize memory allocation strategy for large MIMO systems.       
\newpage
% conference papers do not normally have an appendix
